Something about Performance, 2D Component of NEMO. (TBC)

### 2015-02-24 nemolite2d_perf.tex
More text describing serial results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology(?)}

@@

In Figure~\ref{FIG_orig_perf_summary} we plot the serial performance
of the original version of the nemolite2d code for the range of
compilers and hardware considered here.

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{orig_summary}
\caption{Summary of the performance of the original version of the 
nemolite2d code on 
the range of compilers and hardware under consideration.}
\label{FIG_orig_perf_summary}
\end{figure}

Moving now the to the \psykal version of nemolite2d,
Figure~\ref{FIG_best_psykal_perf_summary} plots the performance of the
fastest \psykal version for each of the compiler/CPU combinations. The
only significant difference between this and the plot of the
performance of the original version of the code in
Figure~\ref{FIG_orig_perf_summary} is that the difference between the
Gnu- and Intel-compiled executables is slightly more marked. This is
clarified in Figure~\ref{FIG_slowdown_summary} which plots the
percentage difference between the performance of the original and the
(best) \psykal versions of nemolite2d for each compiler/CPU combination.

@@

@@

Having show that, in general, we can recover and occasionally improve
upon the performance of the original version of nemolite2d, the next
logical step is to examine the necessary code transformations.  We do
this for the $256^{2}$ case since this fits within cache on both the
Haswell and Ivy Bridge CPUs we are using here.
Table~\ref{TABLE_opt_breakdown} shows detailed timings for this case
after each transformation has been applied to the code. The same data
is visualised in Figure~\ref{FIG_opt_stages_256} where the performance
of the \psykal version for a given compiler/CPU is given relative to
the performance of the original with the same compiler/CPU.

@@

\section{Conclusions}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology(?)}
\label{sec_methodology}

@@

We first consider the performance of the code in serial and examine
the effects of the transformations described in
Section~\ref{sec_methodology}. Once we have arrived at an optimal form
for the serial version of nemolite2d we then investigate its parallel
performance on both CPU- and GPU-based systems.

\subsection{Serial Performance}

In Figure~\ref{FIG_orig_perf_summary} we plot the serial performance
of the original version of the nemolite2d code for the range of
compilers and CPUs considered here. Unlike the Shallow code, the
original version of nemolite2d has not been optimised. Although it is
still a single source file it is, in common with NEMO itself,
logically structured with separate subroutines performing different
parts of the physics within each time step. This structuring and the
heavy use of conditional statements favour the Intel compiler which
significantly out-performs both the Gnu and Cray compilers. Only when
the problem size falls out of cache does the performance gap begin to
narrow. The reason for the performance deficit of the Cray-compiled
binary comes down to SIMD vectorisation, an issue that we explore
below.

\begin{figure}[!t]
\centering
\includegraphics[width=2.8in]{orig_summary}
\caption{Summary of the performance of the original version of the 
nemolite2d code on 
the range of compilers and CPU-based systems under consideration.}
\label{FIG_orig_perf_summary}
\end{figure}

Moving now the to the \psykal version of nemolite2d,
Figure~\ref{FIG_best_psykal_perf_summary} plots the performance of the
fastest \psykal version for each of the compiler/CPU combinations.
While the Intel compiler still produces the best-performing binary,
the Cray-compiled binary is now a very close second. In fact, the
performance of both the Gnu- and Cray-compiled \psykal versions is
generally significantly greater than that of their respective original
versions. We also note that best absolute performance (in terms of
grid points processed per second) with any compiler is obtained with
the $256^2$ domain.

@@

Figure~\ref{FIG_slowdown_summary} plots the percentage difference
between the performance of the original and the best \psykal versions
of nemolite2d for each compiler/CPU combination. This shows that it is
only the Intel-compiled binary running the $64^2$ domain that is
slower with the \psykal version of the code (and then only by some
3\%). For all other points in the space, the optimised \psykal version
of the code performs better. The results for the Cray compiler are
however somewhat sqewed by the fact that it did not SIMD vectorise key
parts of the original version (see below).

@@

Having show that we can recover and often improve upon the performance
of the original version of nemolite2d, the next logical step is to
examine the necessary code transformations in detail.  We do this for
the $256^{2}$ case since this fits within cache on the Ivy Bridge CPUs
we are using here.  Table~\ref{TABLE_opt_breakdown} shows detailed
timings for this case after each transformation has been applied to
the code. The same data is visualised in
Figure~\ref{FIG_opt_stages_256} where the performance of the \psykal
version for a given compiler/CPU is given relative to the performance
of the original with the same compiler/CPU.

@@

\subsection{Parallel Performance}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
